Visual diagnostics for constrained optimisation with application to guided tours

### 2020-10-18 paper/paper.Rmd
Starting to work on Section 3.

@@

Modern development of the line search methods focuses on proposing different computation on the searching direction: $d^{(j)}$ and various approximations on the step size: $\alpha_j$ catered for practical optimisation problems. The specific problem context discussed in this paper is projection pursuit guided tour. 
Projection pursuit and guided tour are two separate methods in exploratory data analysis focusing on different aspects: Projection pursuit, coined by @friedman1974projection, detects interesting structures (i.e. clustering, outliers and skewness) in multivariate data via low dimensions projection whilst guided tour is one particular variation in a broader class of data visualisation method, tour.

Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis. Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection. A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].

As a general visualisation method, A tour produces animated visualisation of the high dimensional data via rotating low dimension planes. Different tours choose the low dimensional planes differently, for example, a grand tour[@cook2008grand] selects the planes randomly in the high dimensional space and manual tour[@cook1997manual] chooses a sequence of bases that gradually phase in and out one variable to view the contribution of that variable in the visualisation. Guided tour, the main interest of this paper, chooses its projection with the aid of projection pursuit. Given a random start, projection pursuit iteratively finds bases with higher index value and guided tour constructs geodesic interpolation between two consecutive to form a tour path. Mathematical details of the geodesic interpolation can be found in @buja2005computational. Figure \@ref(fig:tour-path) shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, the interpolations between two blue frames. The tour method has been implemented in the `tourr` package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].

>>>>

Modern development of line search methods focuses on improving the computation of the searching direction $d^{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems. Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.

@@

The projection pursuit guided tour combines two different methods in exploratory data analysis, focusing on different aspects. Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g. clustering, outliers and skewness) in multivariate data via low dimensions projection. The guided tour is using ideas from projection pursuit to define a particular variation in a broader class of data visualisation methods, building on the grand tour approach [@As85].

To define projection pursuit, we first need to establish the notation used. Let $\mathbf{X}_{n \times p}$ be the data matrix, with $n$ observations in $p$ dimensions. A d-dimensional projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix. Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ (corresponding to an associated projection matrix $\mathbf{A}$) onto an index value $I$. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a projection.

A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].

As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes. Different tour types choose these planes differently, for example, a grand tour[@cook2008grand] selects the planes randomly and a manual tour[@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection. Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit. Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.

Mathematical details of the geodesic interpolation can be found in @buja2005computational. Figure \@ref(fig:tour-path) shows a sketch of the tour path. The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them. The tour method has been implemented in the `tourr` package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
